---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# riceblast

<!-- badges: start -->
<!-- badges: end -->

The goal of `riceblast` is to provides data and tools for forecasting
rice blast disease outbreaks using weather-based parameters. It combines
climate variables (such as temperature, humidity, and rainfall) with
knowledge of disease development to build predictive models and early
warning systems. The package helps researchers, agronomists, and
policymakers understand and anticipate disease risks, supporting
sustainable management strategies to reduce crop losses and improve food
security.

## Installation

You can install the development version of riceblast from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("pridiltal/riceblast")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(riceblast)
head(field1)
```
