---
title: "Blast Detection Framework"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Blast Detection Framework}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(riceblast)
library(ceylon)
library(tidyverse)
library(sf)
library(fable)
library(feasts)
library(tsibble)
library(evd)
library(lubridate)
#library(POT)
library(rlang)
```

This vignette presents a step-by-step weather-based framework for detecting rice blast disease using historical climate records and blast observations from Field 1 in Kilinochchi, Sri Lanka (9.385680° N, 80.538083° E).


## Visualize data

```{r field1, fig.height=12}
# assuming your data frame is named df
# Gather all continuous variables into long format
df_long <- field1 |>
  pivot_longer(
    cols = c(d2m, t2m, stl1, u10, v10, tp, RH),
    names_to = "variable",
    values_to = "value"
  )

# Create the plot
ggplot(df_long, aes(x = time, y = value, color = type)) +
  geom_point(alpha = 0.6) +
  geom_vline(aes(xintercept = as.POSIXct("2021-02-12 00:00:00", tz = "UTC")),
             color = "red", linetype = "dashed") +
  geom_vline(aes(xintercept = as.POSIXct("2021-02-12 00:00:00", tz = "UTC") -
                   (14 * 24 * 60 * 60)), color = "blue", linetype = "dashed") +
  geom_vline(aes(xintercept = as.POSIXct("2020-11-15 00:00:00", tz = "UTC")),
             color = "darkgreen", linetype = "dashed") +
  facet_wrap(~variable, scales = "free_y", ncol = 1) +
  theme_minimal() +
  labs(
    title = "Meteorological Variables Over Time",
    x = "Time",
    y = "Value",
    color = "Type"
  ) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```

We consider the two-and-a-half-month period following the planting date as the *plant’s typical behaviour* window. Previous observations indicate that disease symptoms do not appear within this period. During this window, both the air temperature at 2 meters above ground (t2m) and the top-layer soil temperature (stl1) remain relatively stable. This stability provides an ideal baseline for modelling typical plant behaviour, and we therefore adopt this model to represent the plant’s typical state.

### Air temperature 2 meters above ground

```{r}
# Create the plot of Air temperature 2 meters above ground.
field1 |>
  ggplot(aes(x = time, y = t2m)) +
  geom_line() +
  geom_point(alpha = 0.6) +
  geom_vline(aes(xintercept = as.POSIXct("2021-02-12 00:00:00", tz = "UTC")),
             color = "red", linetype = "dashed") +
  geom_vline(aes(xintercept = as.POSIXct("2021-02-12 00:00:00", tz = "UTC") -
                   (14 * 24 * 60 * 60)), color = "blue", linetype = "dashed") +
  geom_vline(aes(xintercept = as.POSIXct("2020-11-15 00:00:00", tz = "UTC")), 
             color = "darkgreen", linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Air temperature 2 meters above ground",
    x = "Time",
    y = "Value",
    color = "Type"
  ) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```

###  Temperature of the top soil layer.

```{r}
# Create the plot of Air temperature 2 meters above ground.
field1 |>
  ggplot(aes(x = time, y = stl1)) +
  geom_line() +
  geom_point(alpha = 0.6) +
  geom_vline(aes(xintercept = as.POSIXct("2021-02-12 00:00:00", tz = "UTC")), 
             color = "red", linetype = "dashed") +
  geom_vline(aes(xintercept = as.POSIXct("2021-02-12 00:00:00", tz = "UTC") -
                   (14 * 24 * 60 * 60)), color = "blue", linetype = "dashed") +
  geom_vline(aes(xintercept = as.POSIXct("2020-11-15 00:00:00", tz = "UTC")),
             color = "darkgreen", linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Soil temperature level 1",
    x = "Time",
    y = "Value",
    color = "Type"
  ) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```


### Modelling the typical behaviour

```{r}
typical_f1  <- field1 |>
  filter(
    time >= as.POSIXct("2020-11-15 00:00:00", tz = "UTC"),
    time <= as.POSIXct("2020-11-15 00:00:00", tz = "UTC") + months(2) + days(15)
  ) |>
  select(time, t2m) |>
  as_tsibble()
  
typical_f1 |> autoplot() 
```


```{r}
typical_f1 |> 
  gg_tsdisplay(t2m, plot_type='partial')
```

```{r}
 typical_f1 |>
  mutate(diff = difference(t2m, 24)) |>
  ggtime::gg_tsdisplay(diff, plot_type = 'partial')
```





```{r eval=FALSE}

# Step 1: Fit model and compute lower extreme threshold
analysis <- model_extremes(
  full_data = field1,
  time_col = time,
  typical_start = "2020-11-15 00:00:00",  
  typical_end = "2021-02-01 00:00:00",
  response = t2m,
  t_method = "evd"
)

# Step 2: Forecast + test against full data
test_result <- test_extremes(analysis, h = 1000)

# Step 3: Visualize everything
plots <- plot_extreme_analysis(test_result)
```

```{r echo=FALSE}
analysis <- readRDS(system.file("extdata", "analysis_example_evd.rds", package = "riceblast"))
test_result <- readRDS(system.file("extdata", "test_result_example_evd.rds", package = "riceblast"))
plots <- readRDS(system.file("extdata", "plots_example_evd.rds", package = "riceblast"))
```




```{r}
plots$main_plot

```


```{r}
plots$error_plot
```



```{r eval=FALSE}
# Step 1: Fit model and compute lower extreme threshold
analysis <- model_extremes(
  full_data = field1,
  time_col = time,
  typical_start = "2020-11-15 00:00:00",
  typical_end = "2021-02-01 00:00:00",
  response = t2m,
  t_method = "boxplot"
)

# Step 2: Forecast + test against full data
test_result <- test_extremes(analysis, h = 1000)

# Step 3: Visualize everything
plots <- plot_extreme_analysis(test_result)
```

```{r echo=FALSE}
analysis <- readRDS(system.file("extdata", "analysis_example_boxplot.rds", package = "riceblast"))
test_result <- readRDS(system.file("extdata", "test_result_example_boxplot.rds", package = "riceblast"))
plots <- readRDS(system.file("extdata", "plots_example_boxplot.rds", package = "riceblast"))
```


```{r}
plots$main_plot

```



```{r}
plots$error_plot
```


```{r}
analysis$lower_limit

```


