---
title: "visualization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(riceblast)
library(ceylon)
library(tidyverse)
library(sf)
```


## Test

This dataset contains historical climate and rice blast disease observations for Field 1 located in Kilinochchi, Sri Lanka (latitude 9.385680, longitude 80.538083). The data were collected from the Copernicus EU Earth Observation Program (https://www.copernicus.eu/en), which integrates satellite and ground-based information to provide global climate coverage, and supplemented with rice blast disease records obtained from the Department of Agriculture, Sri Lanka. Field visits were conducted during the 2020/2021 “Maha” season when blast outbreaks were reported. Farmers observed that under favorable environmental conditions, complete field infection occurred within 24–48 hours after the appearance of initial symptoms.

```{r}
data(district)
p <- ggplot(district) + geom_sf() 


fields <- tibble::tribble( 
  ~Field,           ~lat,     ~lon,
  "Kilinochchi:Field1",    9.385680, 80.538083
)

points <- st_as_sf(fields, coords = c("lon", "lat"), crs = 4326) |>
  st_transform(crs = st_crs(sf_sl_0))

p <- p + geom_sf(data = points,  colour = "blue", size = 5) +
  ggrepel::geom_label_repel(
    data = points,
    aes(label = Field, geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0,
   colour = "red",
    segment.colour = "red",
   max.overlaps = getOption("ggrepel.max.overlaps", default = 200),
    label.size = 0.04,
   size=4,
   #alpha = 0.7,
   #force_pull = 480,
  force = 70,
   seed = 1
  )

print(p)
```

## Visualize data

```{r field1, fig.height=12}
# assuming your data frame is named df
# Gather all continuous variables into long format
df_long <- field1 |>
  pivot_longer(cols = c(d2m, t2m, stl1, u10, v10, tp, RH),
               names_to = "variable",
               values_to = "value")

# Create the plot
ggplot(df_long, aes(x = time, y = value, color = type)) +
  geom_point(alpha = 0.6) +
  geom_vline( aes(xintercept =  as.POSIXct("2021-02-12 00:00:00", tz = "UTC")), color = "red", linetype = "dashed") +
     geom_vline( aes(xintercept =  as.POSIXct("2021-02-12 00:00:00", tz = "UTC")-(14 * 24 * 60 * 60)), color = "blue", linetype = "dashed") +
       geom_vline( aes(xintercept =  as.POSIXct("2020-11-15 00:00:00", tz = "UTC")), color = "darkgreen", linetype = "dashed") +
  facet_wrap(~ variable, scales = "free_y", ncol = 1) +
  theme_minimal() +
  labs(
    title = "Meteorological Variables Over Time",
    x = "Time",
    y = "Value",
    color = "Type"
  ) +
  theme(
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```

